A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-10-13 paper/paper.tex
Danilo's part quasi-finished

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{mdframed}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{wrapfig}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{ulem}
\usepackage{multirow}

@@

\author{Elvis Dohmatob, (Bert Kappen), Danilo Bzdok\\
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  firstname.lastname@inria.fr}

@@

\begin{abstract}
The default mode network (DMN) is the neural representation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive
envisioning of past and hypothetical scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction.
DMN function is recast in mathematical terms
from the perspective of control theory and
reinforcement learning.
Neurobiological evidence on the human and animal DMN
is linked to recent progress in autonomous systems research
in the machine-learning domain.
We argue that this engineering approach to DMN function can parsimoneously
explain a variety of existing neuroscientific hypotheses.
%
Formal models of the computational processes subserved in the DMN
could offer access routes to statistical optimality
of human behavior.

\end{abstract}

% official NIPS keywords
\textbf{\\keywords}: Systems Biology, mind wandering, cognitive science,
autonomous learning systems
\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
Guillaume Dumas, Smallwood, Tenenbaum

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain inter-neuronal information
transfer in the absence of a focused behavioral goal (Raichle 2001;
Shulman, 1999).
This baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
(Bzdok and Eickhoff, 2016a).
What has early been described as the "stream of consciousness"
in psychology (James, 1890), found its neurobiological manifestation 
in the so-called "default mode network" (DMN)
(Shulman et al., 1997). This networked collection of some of the highest regions in
the human association cortex (Mesulam et al., 1998) consistently
increased in neural activity during
unfocused everyday mind wandering (Raichle et al., 2001). That is, in
the beginning of the 21st century, brain imaging was the first technology
to allow for the discovery of a unique brain network that subserves
baseline mental activity (Bzdok et al., 2012; Bzdok et al., 2015).

The DMN appeared to be exclusive in task-induced
deactivation during various psychological experiments.
This processing of unknown information categories in the DMN has
been argued to mediate an evolutionarily conserved function for the individual.
This is all the more likely because the DMN contains the two
biggest hotspots of energy consumption in the entire central nervous system.
DMN activity also persists to a substantial degree during
sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and future events. This
brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default-mode areas (Weissman et al., 2006)
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts (Mason et al., 2007; Seli et al., 2016). 
%
Understanding of the overarching DMN function is complicated by
simultaneously controlling perception–action cycles and
maintaining baseline contemplations
across time, splace, and content domains
interfacing the external world and the self.

Consider an agent faced by choosing the next action
and being guided by rich
reenactment of really happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
can be the combination of control theory and reinforcement learning.
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\cite{antrobus1966studies, pope1978regulation}.
For instance,
the more the currently executed task is unknown and unpracticed,
the lower the tendency for stimulus-independent thoughts
\cite{filler1973daydreaming, teasdale1995stimulus}.
These \"offline\" processes may however contribute to optimized control of the organism.
We formalize
a policy matrix to capture the space of possible actions that can be performed
on the environment given the current state. A value matrix 
maps maps expected rewards to environmental objects and events.
Switching between states reduces to a sequential processing model.
In a model-based reinforcement learning approach,
we try to understand the environment always better
starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feed-back loops
that are calibrated by error occurrence.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as minimizing prediction error
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
(Tenenbaum2011Science).
\textit{Control} refers to the influences an agent exterts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories
of past experience, random mind-wandering and dreams during sleep
may provide meanginful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.

@@

The left TPJ in turn exhibits a close topographical relationship to
Wernicke's area
involved in the comprehension or understanding of written and spoken language.
Neurological patients with damage caused to Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Their speech
preserves natural rhythm and about normal syntax, yet the
voiced sentences are devoid of meaning (i.e., aphasia).
Abstracting from the typical semantic interpretations in linguistics
and neuropsychology,
the LTPJ probably mediates access to and integrating of world knowlege
to model action concepts
\cite{binder2011neurobiology, seghier2013angular}.
For instance, LTPJ lesions also entail problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing any language content
(Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Also inner speech hinges on knowledge retrieval of statistical structure
about the physical and inter-personal world.
Multivariate volumetric analyses showed that the internal production of
formulated thought ("language of the mind") is closely related to the left TPJ
(Geva et al., 2011).
A number of neuroimaging studies demonstrated LTPJ activity responses
to sementically aberrant words in sentences, which suggests a role
in binding a given semantic concept into an integrated whole.
Further,
episodic memory recall and imagination strongly draw on
complex world knowledge.
Isolated building blocks of world statistics probably get reassembled
in internally generated visual scenarios that
navigate present action, weigh hypothetical possibilities, and forcast the future.
%
The LTPJ may hence facilitate the automated prediction of events
by incorporating experience-derived models of the world
into ongoing action, planning, and problem solving.
world -> policy matrix driven computation
self -> value matrix driven computation

@@

\subsection{Correspondence between DMN function and RL framework}
\textbf{XXX Put the diagrams and process charts here.}
\begin{itemize}
  \item The policy matrix represents the repertoire of possible actions
  on the world given a current state. It encodes the probabilites of
  choosing actions to be executed in a certain situation.
  \item Interactions between external and internal sources of
  information are probably commonplace during perception
  \item The MDP frameworks how the human brain could solve
  environmental problems over extended time periods.
  \item Given a state, the value function represents the estimated utility of
  being in that state, which is subjective to the agent. 
  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
  "How desirable is it to be in state $s$ ?"
  For each action $a$ that can be taken from state $s$,
  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
  utility of executing a specific action; it answers the question
  "What is the expected utiltiy of taking action $a$ in this situation ?"
  \item The DMN may subserve
  constant exploration of possible future actions and their
  cumulative outcomes. Implicit computation of future choices 
  provides an explanation for purposeful mind-wandering.
  \item hippocampus -> provides perturbed action-transition-state-reward samples
  as a block of "imagined", "hypothesized", "recalled" experience;
  the small variations in these experience blocks allow searching
  a larger space of parameters and possible experiences.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep) pseudo-experiencing possible
  future scenarios and action outcomes.
  Our approach acknowledges the unavoidable stochasticity of
  computation in neuronal processes \cite{faisal2008noise}.
  \item
  \textit{Inference} in the human brain reduces to generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  \item In the RL view, plasticity in the DMN arises naturally.
  If an agent behaving optimally in a given environment moves
  to novel, yet unexperienced environments, reward prediction errors will
  massively increase. This will lead to adaptation of the policy matrix
  and value function until the system converges to a
  new steady-state of optimal action decisions given certain states.
  \item The DMN is today known to consistently increase in neural
  activity when engaging cognitive processes that are detached from
  current sensory environment. DMN involvement in thinking about the past,
  the future, and hypothetical possibilities ties in with the
  MDP's implicit computation of action and state cascades as a function
  of what happened in the past.

\end{itemize}

@@

\section{Relation to other cognitive models}
\subsection{Predictive Coding Hypothesis}
The predictive coding and Bayesian brain framework 
is a frequently evoked idea related to the default mode function
\cite{bar2007}.
According to this functional account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information. 
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact,
back projections probably outnumber by far the existing input projection in the brain
\cite{salin1995corticocortical}.
This can explains the constructive, generative nature of sensory perception
\cite{friston2010free} and
why motor action is intimately related to sensory expectations
\cite{wolpert1995internal}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back-projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how data arise in the
world would hence be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the incoming sensory examples.

@@

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
functional account that was proposed for the DMN \cite{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflects the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation
\cite{}.
-> the most behaviorally relevant stimuli
-> particularly in the secur- ing of mates, protection, and resources
-> tracking for salient and unexpected events

This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-relevant information processing

Environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\cite{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems in a social way
\cite{whiten1988machiavellian}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecific
turned into a decisive evolutionary advantage.

-> contents of memory and everyday dreaming tied to people and relationships to them

Cognitive neuroimaging studies linked the DMN to the contemplation of others’ (Mar, 2011) and one’s thoughts (Lombardo et al., 2009) mental states, 
Schilbach et al., 2008
Clan mentality

role increasing during self-referential tasks.

 \begin{itemize}
    \item DMN does social cognition -> but it exists in rats
    and monkeys with much more poorly developped social capacities
    -> our computational framework can be equally applied to humans
    and other primates.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explorated and which decision trees are neglected,
    which implies allocation of attention and computational load.
    \item How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance? Yes, if
    environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set? -> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
 \end{itemize}

\section{Relation to other statistical models (Elvis)}
\paragraph{Control theory.}
\begin{itemize}
  \item The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
\end{itemize}

\paragraph{Reinforcement Learning.}
abc

\paragraph{Free Energy Principle.}
\begin{itemize}
  \item Closely related to Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  \item A common point between the Free Energy principle and our framework
  is placing the minimization of suprise at the core of brain function.
  In RL, surprise minimization reduces to accurate prediction of 
  rewarding outcomes in the future.
  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute the free energy principle
in practice, which makes it physiologically less plausible.
  \end{itemize}

\section{Conclusion}
What brain function could be important enough
for the existence and survival of the human species
to warrant constant, high energy costs?
MDPs provide an attractive
computational model how the human association cortex
might implement supra-modal control of the environment by
obtaining reward and avoiding pain to
optimize the organism's fate.
This idealized process model explains
a number of previous experimental observations in the
DMN literature by a simple but non-trivial concept.
%
Viewed as an MDP, behavior unfolds by integrating happened past events
and possible future events to guide action choice in the present context.
This functional account is more compatible with the DMN's
poorly understood involvement across
autobiographical memory retrieval, problem solving,
abstract reasoning based on internally generated scenes, social cognition,
as well as delay discounting and self-related prospection.
MDPs provide a mathematical formalism how
optimal substructure can be recursively exploited
when confronted with nested decisions.
Algorithmic gains by injecting stochasticity in the recall of past
behavior and outcomes may plausibilize why
very accurate memories have been evolutionarily disfavored
and why human creativity might be adaptive.
%
In principle,
neuroscientific experiments can be designed that operationalize
the set of action, value, and state variables that determine
the behavior of intelligent RL systems.
The proposed machine-learning
approach to DMN function is hence not only practically computable but
also neuroscientifically falsifiable.
At the least, we propose an alternative vocabulary to
describe and interpret neuroscientific studies in DMN research.
%
Ultimately,
the DMN can be viewed as a smoothing kernel extending
from the relatively recent events to potentially upcoming events
to constantly improve sensory processing and
action performance in the world.

@@

\small
\bibliographystyle{splncs03}
\bibliography{refs}

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage[round]{natbib}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{mdframed}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{wrapfig}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{ulem}
\usepackage{multirow}

@@

\author{Elvis Dohmatob, Bert Kappen, Danilo Bzdok\\
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  firstname.lastname@inria.fr}

@@

\begin{abstract}
The default mode network (DMN) is the neural representation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive
envisioning of past and hypothetical scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction.
DMN function is recast in mathematical terms
from the perspective of control theory and
reinforcement learning.
In this way,
neurobiological evidence on the human and animal DMN
is linked to recent progress in autonomous systems research
and machine-learning.
We argue that this engineering approach to DMN function can parsimoneously
explain a variety of existing neuroscientific hypotheses.
%
Formal models of the computational processes subserved in the DMN
could offer access routes to statistical optimality
of human behavior.
\end{abstract}

% official NIPS keywords
\textbf{\\keywords}: systems biology, mind wandering, cognitive science,
autonomous learning systems
% \textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
% Guillaume Dumas, Smallwood, Tenenbaum

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
potentially purposeful neuronal computation
in the absence of a focused behavioral goal
\cite{kenet2003spontaneously, fiser2004small}.
This baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
(Bzdok and Eickhoff, 2016a).
What has early been described as the "stream of consciousness"
in psychology (James, 1890), found its neurobiological manifestation 
in the so-called "default mode network" (DMN)
(Shulman et al., 1997). This networked collection of some of the highest regions in
the human association cortex (Mesulam et al., 1998) consistently
increased in neural activity during
unfocused everyday mind wandering (Raichle et al., 2001). That is, in
the beginning of the 21st century, brain imaging was the first technology
to allow for the discovery of a unique brain network that subserves
baseline mental activity (Bzdok et al., 2012; Bzdok et al., 2015).

The DMN appeared to be exclusive in task-induced
deactivation during various psychological experiments.
This processing of unknown information categories in the DMN has
been argued to mediate an evolutionarily conserved function for the individual.
This is all the more likely because the DMN contains the two
biggest hotspots of energy consumption in the entire central nervous system.
DMN activity also persists to a substantial degree during
sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and future events. This
brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default-mode areas (Weissman et al., 2006)
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts (Mason et al., 2007; Seli et al., 2016). 
%
Understanding of the overarching DMN function is complicated by
simultaneously controlling perception–action cycles and
maintaining baseline contemplations
across time, splace, and content domains
at the interface of the external world and the self.

Consider an agent faced by the choice of the next action
and being guided by rich
reenactment of really happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
can be the combination of control theory and reinforcement learning.
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\cite{antrobus1966studies, pope1978regulation}.
For instance,
the more the currently executed task is unknown and unpracticed,
the lower the tendency for stimulus-independent thoughts
\cite{filler1973daydreaming, teasdale1995stimulus}.
These ``offline'' processes may however contribute to optimized control of the organism.
We formalize
a policy matrix to capture the space of possible actions that can be performed
on the environment given the current state. A value matrix 
maps expected rewards to environmental objects and events.
Switching between states reduces to a sequential processing model.
In a model-based reinforcement learning approach,
we try to understand the environment always better
starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feed-back loops
that are calibrated by prediction error.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\cite{tenenbaum2011grow}.
\textit{Control} refers to the influences an agent exterts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories
of past experience, random mind-wandering and dreams during sleep
may provide meanginful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\cite{raichle2005intrinsic}.
This paper proposes a formal model that
satisfies this contention.

@@

The left TPJ in turn exhibits a close topographical relationship to
Wernicke's area
involved in the comprehension or understanding of written and spoken language.
Neurological patients with damage caused to Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Their speech
preserves natural rhythm and about normal syntax, yet the
voiced sentences are devoid of meaning (i.e., aphasia).
Abstracting from the typical semantic interpretations in linguistics
and neuropsychology,
the LTPJ probably mediates access to and integrating of world knowlege
to model action concepts
\cite{binder2011neurobiology, seghier2013angular}.
For instance, LTPJ lesions also entail problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing any language content
(Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Also inner speech hinges on knowledge retrieval of statistical structure
about the physical and inter-personal world.
Multivariate volumetric analyses showed that the internal production of
formulated thought ("language of the mind") is closely related to the left TPJ
(Geva et al., 2011).
A number of neuroimaging studies demonstrated LTPJ activity responses
to sementically aberrant words in sentences, which suggests a role
in binding a given semantic concept into an integrated whole.
Further,
episodic memory recall and imagination strongly draw on
complex world knowledge.
Isolated building blocks of world statistics probably get reassembled
in internally generated visual scenarios that
navigate present action, weigh hypothetical possibilities, and forcast the future.
%
The LTPJ may hence facilitate the automated prediction of events
by incorporating experience-derived models of the world
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

@@

\subsection{Correspondence between DMN function and RL framework}
\textbf{XXX Put the diagrams and process charts here.}
\begin{itemize}
  \item The policy matrix represents the repertoire of possible actions
  on the world given a current state. It encodes the probabilites of
  choosing actions to be executed in a certain situation.
  \item Interactions between external and internal sources of
  information are probably commonplace during perception
  \item The MDP frameworks how the human brain could solve
  environmental problems over extended time periods.
  \item Given a state, the value function represents the estimated utility of
  being in that state, which is subjective to the agent. 
  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
  "How desirable is it to be in state $s$ ?"
  For each action $a$ that can be taken from state $s$,
  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
  utility of executing a specific action; it answers the question
  "What is the expected utiltiy of taking action $a$ in this situation ?"
  \item The DMN may subserve
  constant exploration of possible future actions and their
  cumulative outcomes. Implicit computation of future choices 
  provides an explanation for purposeful mind-wandering.
  \item hippocampus -> provides perturbed action-transition-state-reward samples
  as a block of "imagined", "hypothesized", "recalled" experience;
  the small variations in these experience blocks allow searching
  a larger space of parameters and possible experiences.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep) pseudo-experiencing possible
  future scenarios and action outcomes.
  Our approach acknowledges the unavoidable stochasticity of
  computation in neuronal processes \cite{faisal2008noise}.
  \item
  \textit{Inference} in the human brain reduces to generalization of
  policy and value adaptations from sampled experiences to
  successful action choices and reward predictions in future states.
  \item In the RL view, plasticity in the DMN arises naturally.
  If an agent behaving optimally in a given environment moves
  to novel, yet unexperienced environments, reward prediction errors will
  massively increase. This will lead to adaptation of the policy matrix
  and value function until the system converges to a
  new steady-state of optimal action decisions given certain states.
  \item The DMN is today known to consistently increase in neural
  activity when engaging cognitive processes that are detached from
  current sensory environment. DMN involvement in thinking about the past,
  the future, and hypothetical possibilities ties in with the
  MDP's implicit computation of action and state cascades as a function
  of what happened in the past.
\end{itemize}

@@

\section{Relation to other cognitive models}
\subsection{Predictive Coding Hypothesis}
The predictive coding and Bayesian brain framework 
is a frequently evoked idea related to the default mode function
\cite{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information. 
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact,
back projections probably outnumber by far the existing input projection in the brain
\cite{salin1995corticocortical}.
This can explains the constructive, generative nature of sensory perception
\cite{friston2010free} and
why motor action is intimately related to sensory expectations
\cite{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back-projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how data arise in the
world would hence be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the incoming sensory examples.

@@

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \cite{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflects the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain's baseline realizes
a ``radar'' function to
detect subjectively salient and unexpected events that
are the most important for unfolding behavior,
including resources,
the securing of mates, and protection.
%
In fact,
environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\cite{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social connections
\cite{whiten1988machiavellian}.
Highly efficient neurobiological processing of social cues is exemplified by
facial judgments being routinely processed in less than 100 ms
\cite{bar2006very}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecifics
turned into a decisive evolutionary advantage
\cite{frith2010social}.
%
This may explain why social topics amount to roughly
two thirds of human everyday communication \cite{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and the complex relationships between them.

 \begin{itemize}
    \item Processing social information has been proposed to underlie the
    physiological baseline of human brain function
    \cite{schilbach2008minds}. This was later challenged by observing
    analogues of the DMN in monkeys \cite{mantini2011default}
    and rats \cite{lu2012rat}, with
    supposedly less advanced social capacities.
    The MDPs parsimoneously explain the dominance of social content in
    human mental scenes by their extremely high relevance to humans
    reflected by high values for social information in the value matrix.
    \item How can a same neurobiological circuit be equally important
    for baseline house-keeping functions and specific task performance?
    Our computational account can explain why the DMN is implicated
    in both a goal-directed task and an idlying rest cognitive set,
    if environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    In tasks, the policy and value matrices are updated to optmize short-term action,
    whereas, at rest, these parameter updates may
    serve especially mid- and long-term action.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explored by the agend and which decision trees are neglected.
    This directly implies allocation of attention and computational load.
    \item In sum,
    MDPs imply social thinking as a physiological brain baseline
    as a special case, without excluding other topics.
 \end{itemize}

\section{Relation to other statistical models (Elvis)}
\subsection{Control theory.}
\begin{itemize}
  \item The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
\end{itemize}

\subsection{Reinforcement Learning.}
abc

\subsection{Free Energy Principle.}
\begin{itemize}
  \item Closely related to Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  \item A common point between the Free Energy principle and our framework
  is placing the minimization of suprise at the core of brain function.
  In RL, surprise minimization reduces to accurate prediction of 
  rewarding outcomes in the future.
  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute the free energy principle
in practice, which makes it physiologically less plausible.
  \end{itemize}

\section{Conclusion}
What brain function could be important enough
for the existence and survival of the human species
to warrant constant, high energy costs?
MDPs provide an attractive
computational model how the human association cortex
might implement supra-modal representation and control of the environment to
optimize the organism's fate by
obtaining reward and avoiding pain.
This idealized process model explains
a number of previous experimental observations in the
DMN literature by a simple but non-trivial mechanism.
%
Viewed as an MDP, behavior unfolds by integrating happened past events
and possible future events to guide action choice in the present context.
This functional account is more compatible with the DMN's
poorly understood involvement across
autobiographical memory retrieval, problem solving,
abstract reasoning based on internally generated scenes, social cognition,
as well as delay discounting and self-related prospection.
MDPs provide a mathematical formalism how
optimal substructure in the environment can be recursively exploited
when confronted with nested decisions.
Algorithmic gains by injecting stochasticity in the recall of past
behavior and outcomes to improve the internal world model may explain why
very accurate memories have been evolutionarily disfavored
and why human creativity might be adaptive.
%
In principle,
neuroscientific experiments can be designed that operationalize
the set of action, value, and state variables that determine
the behavior of intelligent RL systems.
The proposed machine-learning
approach to DMN function is hence not only practically computable but
also neuroscientifically falsifiable.
At the least, we propose an alternative vocabulary to
describe and interpret neuroscientific studies in DMN research.
%
Ultimately,
the DMN can be viewed as a smoothing kernel extending
from the relatively recent events to anticipated upcoming events
to constantly improve sensory processing and
action performance in the world.

@@

\small
% \bibliographystyle{splncs03}
% \bibliographystyle{natbib}
\bibliographystyle{plainnat}
\bibliography{refs}
